---
title: "LCMS package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LCMS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette aims to describe how to analyse semi-targeted data in lipid analysis. 

### Preliminar remarks

The data should be obtained as .raw files. These data should be converted in .mzXML
(via MSConvert for example). 

The analysis requires three imputs: 

* the .mzXML files stemming from the raw data
* an integrationTable containing all the species to be targeted
*  a classTable indicating for each class the minimal and maximal retention time


Note: the integrationTable has to vary according to the mode of acquisition (positive or negative). It contains several columns: name (name of the species); compo (composition); class ( lipid class) and mz (theoretical m/z ratio to be detected)

The objective of semi-targeted analysis is to automatize the calculation of the intensity of a list of molecules (in integrationTable)
A directory where all the produced files will be saved is mandatory. 

### Entering parameters of interest

The following parameters should be entered
 
``` {r, echo =TRUE, eval=FALSE}
repoData="E:/ROCOR/Meibum_221108/xml/NEG" # repository containing the .mzXML files
reposave="E:/ROCOR/Meibum_221108/dataAnalysis/NEG" # repository where the results are saved
centroided=FALSE #are the data acquired in centroid mode ?
ppmThreshold=ppm=15 # ppm limit
files=list.files(repoData) # list of file names to be analyzed
integrationTablePath="E:/ROCOR/Meibum_221108/20221108_IonsTable_Neg.xlsx"
classTablePath="E:/ROCOR/Meibum_221108/20221108_RT_Neg.xlsx"
nameFileExcel="excelResults.xlsx" # name of the file containing the results
output="int" # could also be "pct"
```

### Reading the integration table

The integration table and class table are red
``` {r, echo = TRUE,eval=FALSE}
integrationTable=read.xlsx(integrationTablePath)
colnames(integrationTable)=c("class","name","compo","mz")
classTable=read.xlsx(classTablePath)
``` 

### Checking the integration table

This command allows to check that no duplicate exist in the integrationTable

``` {r, echo = TRUE,eval=FALSE}
checkingIntegrationTable(integrationTable,duplicatedName="name")
```

### Checking the class table

 Are the class table and the integration table containing the same class names ?
 
``` {r, echo = TRUE,eval=FALSE}
checkingClassTable(integrationTable=integrationTable,classTable=classTable)
```

### Calculating the intensities in R

 Let's calculate the intensity for all files in R. The result is a list of tables containing all intensities.
 
``` {r, echo = TRUE,eval=FALSE}
res=calculateIntensity(repoData=repoData,files=files,integrationTable=integrationTable,classTable=classTable,centroided=centroided,ppmThreshold = ppmThreshold,ppm=ppm)
```

### Saving all these files in .csv files

Let's save same in a specific repository

``` {r, echo = TRUE,eval=FALSE}
getCsvOfIntensity(res, reposave=reposave)
```

### Compiling all the .csv files to obtain a final excel file of results
 Let's obtain an excel file with all the results
``` {r, echo = TRUE,eval=FALSE}
excel=getExcelOfIntensity(reposave,nameFile=nameFileExcel,output=output)
```